---
title: "MidtermIhorMarkevych"
author: "Ihor Markevych"
date: "10/26/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: 
      collapsed: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
```

# Task

The subspace W is defined as the span of $u_1$ and $u_2$ ($W$ = Span{$u_1$, $u_2$}).
The coordinate of a point $y$ (which doesn’t lie on $W$) is given by the vector $y = \begin{bmatrix} -1 \\ -5 \\ 10 \end{bmatrix}$. Find the shortest distance from $y$ to  $W$ if $u_1 = \begin{bmatrix} 5 \\ -2 \\ 1 \end{bmatrix}$ and $u_2 = \begin{bmatrix} 1 \\ 2 \\ -1 \end{bmatrix}$.


# Functions

## Dot
```{r}
dot <- function(v, w){
    sum(v * w)
}
```

## Norm

```{r}
norm <- function(v){
    sqrt(dot(v, v))
}
```

## Project

```{r}
# project u on v
project <- function(u, v){
   dot(u, v) / dot(v, v) * v
}
```

# Vectors input

```{r}
u1 <- c(5, -2, 1)
u2 <- c(1, 2, -1)

y <- c(-1, -5, 10)
```

# Solution

Shortest distance between vector and a spane is a norm (length) of a orthogonal part of the vector.  
Othrogonal part of the vector is defined as $orth_W(y) = y - proj_W(y)$, where $proj_W(y)$ is a projection of y on spane W.  
For our spane $proj_W(y) = proj_W(u_1) + proj_W(u_2)$.

```{r}
norm(y - project(y, u1) - project(y, u2))
```
