---
title: "pageRank"
author: "Ihor Markevych"
date: "10/25/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(igraph)
```

# Functions

## Get transition matrix from graph

```{r}
getTransMatr <- function(g){
  A <- t(as_adj(g, sparse=F))
  A <- sweep(A, 2, colSums(A), `/`)
  A[is.nan(A)] <- 0
  return(A)
}
```

## Get google matrix

```{r}
getGoogleMatr <- function(A, p=.15){
  S <- A
  S[, getDanglingNodes(S)] <- rep(1 / nrow(A), nrow(A))
  G <- (1 - p) * S + p * matrix(rep(1 / nrow(A), nrow(A) ** 2), nrow = nrow(A))
  return(G)
}
```

## Get dangling nodes

```{r}
getDanglingNodes <- function(A){
  danglingNodes <- c()
  for(i in nrow(A)){
    if(A[i,i] == 1){
      danglingNodes <- c(danglingNodes, i)
    }
  }
  return(danglingNodes)
}
```

## Check graph connectivity

```{r}
getComponentsFromGraph <- function(g){
  components = list()
  
  for(i in unique(components(g)$membership)){
      components = append(components, list(which(components(g)$membership == i)))
  }
  
  return(components)
}
```

## Get Page Rank

```{r}
getPageRank <- function(A){
  G <- getGoogleMatr(A)
  vec <- Re(eigen(G)$vectors[,1])
  return(vec / sum(vec))  
}
```

# Graph 1

```{r}
g1 <- graph( c(1,2, 1,3, 1,4, 2,3, 2,4, 3,1, 4,3, 4,1 ) )

plot(g1)

get.adjacency(g1)
```

##  Transition matrix

### Manually

$$ \begin{bmatrix}
0 & 0 & 1 & \frac{1}{2} \\
\frac{1}{3} & 0 & 0 & 0 \\
\frac{1}{3} & \frac{1}{2} & 0 & \frac{1}{2} \\
\frac{1}{3} & \frac{1}{2} & 0 & 0 \\
\end{bmatrix}$$

### By code

```{r}
A1 <- getTransMatr(g1)
A1
```

## Google Matrix

```{r}
getGoogleMatr(A1)
```

##  Dangling node

Graph 1 does not have dangling node.

```{r}
if(length(getDanglingNodes(A1)) > 0){
  print(getDanglingNodes(A1))
} else {
  cat('No dangling nodes.')
}
```

## Disconnected components

```{r}
getComponentsFromGraph(g1)
```

As it can be seen, graph is connected (no disconnected components).


## Page Rank

```{r}
getPageRank(A1)
```

So, rank should be next: 

1. Page 1.
1. Page 3.
1. Page 4.
1. Page 2.

# Graph 2

```{r}
g2 <- graph( c(1,2, 1,3,  1,4, 2,4, 2,5, 3,1, 4,2, 4,3, 5,5) )

plot(g2)
```

##  Transition matrix

### Manually

$$ \begin{bmatrix}
0 & 0 & 1 & 0 & 0 \\
\frac{1}{3} & 0 & 0 & \frac{1}{2} & 0 \\
\frac{1}{3} & 0 & 0 & \frac{1}{2} & 0 \\
\frac{1}{3} & \frac{1}{2} & 0 & 0 & 0 \\
0 & \frac{1}{2} & 0 & 0 & 1 \\
\end{bmatrix}$$

### By code

```{r}
A2 <- getTransMatr(g2)
A2
```

## Google Matrix

```{r}
getGoogleMatr(A2)
```

##  Dangling node

Graph 2 has dangling node. This is node 5.

```{r}
if(length(getDanglingNodes(A2)) > 0){
  print(getDanglingNodes(A2))
} else {
  cat('No dangling nodes.')
}
```

## Disconnected components

```{r}
getComponentsFromGraph(g2)
```

As it can be seen, graph is connected (no disconnected components).


## Page Rank

```{r}
getPageRank(A2)
```

So, rank should be next: 

1. Page 1.
1. Page 2, 3, 4.
1. Page 5.