---
title: "PCA"
author: "Ihor Markevych"
date: "11/11/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Function 

## Center

```{r}
center <- function(d){
  mean <-  colMeans(d)
  dCentered <- sweep(d, 2, mean, '-')
  dCentered
}
```

## Covariance matrix

```{r}
covarianceMatrix <- function(d){
  cov <- t(d) %*% d / (ncol(d) - 1)
  cov  
}
```

## PCA

**Input:**  
* `d` - a data frame that contains the data set  
* `desiredVariance` - a scaler that denotes the desired percentage of variance preserved in the transformed data set

**Output:**  
* `number` - the minimum number of principal components that is necessary to retain the desired variance specified by the user (`desiredVariance`)  
* `PCA` - the principal components corresponding to previous part, in a matrix   
* `transformed` - the transformed data set in a matrix

```{r}
myPCA <- function(d, desiredVariance=.95){
  d <- data.matrix(d)
  eigens <- eigen(covarianceMatrix(center(d)))
  number <- which(cumsum(eigens$values / sum(eigens$values)) >= desiredVariance)[1]
  
  eigens$vectors <- apply(eigens$vectors, 2, function(x) x / sqrt(sum(x ^ 2)))
  
  PCA <- eigens$vectors[,1:number]
  
  transformed <- d %*% PCA
  
  list(number, PCA, transformed)
}
```

# Test function

Test created function on the first `5` features of R `longley` data set with the desired retained variance of `0.98`. Longley data set is a data frame with 7 features related to economics. You will use the following features for this test: `GNP.deflator`, `GNP`, `Unemployed`, `Armed.Forces`, and `Population`. 

```{r}
myPCA(longley[,1:5], .98)
```