---
title: "MidtermIhorMarkevych"
author: "Ihor Markevych"
date: "10/26/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: 
      collapsed: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
```

# Task

Numerous factors affect cars’ gasoline mileage. In this problem we’ll consider the relationship between gasoline mileage (mpg) and engine displacement (liters).  
  
1. Use the data in `mileage.csv`to
create a linear regression model $mileage = \beta_0 + \beta_1 \frac{1}{\sqrt{disp}}$. Report the model parameters.
1. Use your model to predict the mpg of a car whose engine
displacement is 1.8 liters.


# Functions

## Dot
```{r}
dot <- function(v, w){
    sum(v * w)
}
```

## Norm

```{r}
norm <- function(v){
    sqrt(dot(v, v))
}
```

## Least Squares Solution

Next function computes $\hat{x}$ for Least Squares Model.

```{r}
xHat <- function(A, b) {
  inv(t(A) %*% A) %*% t(A) %*% b
}
```

## Least Squares Error

Next function computes error of Least Squares model.

```{r}
lSError <- function(A, x, b) {
  norm(A %*% x - b)
}
```

## Create Model

Next function will create Least Squares model from dataframe.

```{r}
# d - data, xCols - vector of column names of predictors, y - name of goal column
leastSquares <- function(d, xCols, yCol) {
  y <- as.matrix(d[yCol])
  X <- as.matrix(cbind(replicate(length(y), 1), d[xCols]))
  sol <- xHat(X, y)
  return(c(lSError(X, sol, y), sol))
}
```

## Print Model and Error

Let's create function for printing model formula and its error.

```{r}
printModel <- function(coefs, goal, modelPredictors, error) {
  cat(goal, ' = ', coefs[1])
  for(i in 1:length(modelPredictors)){
    cat(' + ', coefs[i + 1], ' * ', modelPredictors[i])  
  }
  cat('\n')
  cat('Error of this model is ', error, '.\n')
}  
```

## Predict using model

Next function will predict goal variable using vector of coefficients and predictors values.

```{r}
predict <- function(coefs, values){
  sum <- coefs[1]
  for(i in 1:length(values)){
    sum <- sum + coefs[i + 1] * values[i]
  }
  return(sum)
}
```

# Data reading

```{r}
d <- read.csv('mileage.csv')
```

# Data adaptation

```{r}
d['adaptDisp'] <- 1 / sqrt(d['DISP'])
```

# Model

```{r}
goalCol <- 'MILEAGE'

modelPredictors <- 'adaptDisp'

model <- leastSquares(d, modelPredictors , goalCol)
modelErr <- model[1]
model <- model[-1]
```

## Coefficients

$\beta_0$ = `r model[1]`; 
$\beta_1$ = `r model[2]`. 

## Error

Error of this models is `r modelErr`.

## Formula

General formula and error:

```{r}
printModel(model, goalCol, '1 / sqrt(displacement)', modelErr)
```

# Prediction

Prediction of Mileage based on displacement of 1.8:

```{r}
dispToPredict <- 1.8
adaptDispToPredict <- 1 / sqrt(dispToPredict)

cat('Predicted mileage is', predict(model, adaptDispToPredict), 'mpg.')
```
